---
title: "forecasting ED"
author: "Janice Hsu"
date: "2023-08-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fpp3)
library(hts)
library(lubridate)
```


```{r}
data <- read.csv("HLTH0037_ts_cleaned.csv")
```

```{r}
data %>%
  select(Hospital_Hierarchy, Organisation) %>%
unique() 
```

```{r}
data1 <- data %>%
  mutate(YearMonth = ym(YearMonth)) |> 
  mutate(YearMonth = yearmonth(YearMonth)) %>%
  as_tsibble(index = YearMonth, key = c(Age_Code, Sex_ItemName_ENG, Hospital_Code, Hospital_ItemName_ENG)) 

data1 <- data1 %>%
  mutate(Number = 1)# what is this number?
```


```{r}
#Produce a table or plot to show the hierarchy between the organisation and hospital hierarchy

data2 <- data1 %>%
  select(YearMonth, Hospital_ItemName_ENG, Hospital_Hierarchy, Organisation, Number)
```


```{r}
# Convert to data.table
library(data.table)
setDT(data2)

# Create a hierarchical table using data.table operations
hierarchical_table <- data2[, .(Total = sum(Number)), by = .( Organisation, Hospital_ItemName_ENG)]

knitr::kable(hierarchical_table)

```


```{r}
#Number of patients entering ED under different hospital hierarchy

data1_hts <- data1 %>%
  aggregate_key(Organisation/Hospital_ItemName_ENG, Number = sum(Number))
  
data1_hts |>
  filter(is_aggregated(Hospital_ItemName_ENG)) |>
  autoplot(Number) +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```


•	A couple of Local Health Boards (LHBs) were redefined from the 1st of April 2019 onwards: Cwm Taf (27)--> Cwm Taf Morgannwg (30)// Abertawe Bro Morgannwg (26) --> Swansea Bay (31). Therefore, if you decide to forecast at LHB resolution, you might want to consider these 4 as a unique one. 
•	A the Princess of Wales Hospital changed its Local Health Boards
•	So we analyse these 4 as one organisation


```{r}

data1_grouped <- data1 %>%
  mutate(Grouped_Organisation = case_when(
    Organisation %in% c("Cwm Taf", "Cwm Taf Morgannwg", "Abertawe Bro Morgannwg", "Swansea Bay") ~ "Grouped_organisation",
    TRUE ~ Organisation
  ))

```


```{r}
unique(data1_grouped$Grouped_Organisation)
```


```{r}
data2_hts <- data1_grouped %>%
  group_by(Grouped_Organisation) %>%
  summarise(Number = sum(Number))
```


```{r}
data2_hts |>
  filter(!is_aggregated(Hospital_Hierarchy)) |>
  ggplot(aes(x = YearMonth, y = Number)) +
  geom_line(stat = "identity") +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Grouped_Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


```{r}
library(gt)

data2_hts %>%
  select(YearMonth, Number, Grouped_Organisation) %>%
  gt() %>%
  tab_header(title = "Number of patients who enter ED") %>%
  cols_label(
    YearMonth = "Year/Month",
    Number = "Number of Patients",
    Grouped_Organisation = "Organisation Group"
  )

```

```{r}
unique(data1_grouped$Age_Code)
```


Age group: "0-4", "5-17", "18-69", "70^"
```{r}
data1_grouped_age <- data1_grouped %>%
  filter(Age_Code != "Unknown") %>%
  mutate(Grouped_Age = case_when(
    Age_Code == "0 to 4" ~ "0-4",
    Age_Code == "5 to 17" ~ "5-17",
    Age_Code %in% c("18 to 24", "25 to 29", "30 to 34", "35 to 39", 
                    "40 to 44", "45 to 49", "50 to 54", "55 to 59", 
                    "60 to 64", "65 to 69") ~ "18-69",
    Age_Code %in% c("70 to 74", "75 to 79", "80 to 84", "85") ~ "70 and over",
    TRUE ~ "Other"
  ))

```


```{r}

data1_gts <- data1_grouped_age %>%
  filter(!Sex_ItemName_ENG == "Not Specified  or invalid") %>%
  aggregate_key(Grouped_Age* Sex_ItemName_ENG , Number = sum(Number))

data1_gts |>
  filter(!is_aggregated(Sex_ItemName_ENG), !is_aggregated(Grouped_Age)) |>
  autoplot(Number) +
  labs(y = "Number of patients")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```





```{r}
# Number of patients entering ED, facet by sex

data3_ghts <- data1 %>%
  aggregate_key((Organisation/Hospital_Hierarchy)* Sex_ItemName_ENG , Number = sum(Number))

data3_ghts |>
  filter(!is_aggregated(Hospital_Hierarchy), !is_aggregated(Hospital_Hierarchy)) |>
  autoplot(Number) +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Sex_ItemName_ENG), scales = "free_y", ncol = 3) +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```



```{r}
#forecast
data1_full <- data1_grouped_age |>
  aggregate_key((Grouped_Organisation/Hospital_ItemName_ENG) * Grouped_Age, Number = sum(Number))

data1_full <- data1_full |> fill_gaps()
data1_full <- data1_full %>%
               tidyr::drop_na(Number)



fit <- data1_full |>
  filter(year(YearMonth) < 2023) %>%
  model(base = ETS(Number)) |>
  reconcile(
    bu = bottom_up(base),
    ols = min_trace(base, method = "ols")
  )
```


```{r}
fc <- fit |> forecast(h = 6)
```

```{r}
colSums(is.na(fc))
```

```{r}
fit_bu <- fit |> reconcile(bu = bottom_up(base))
fit_ols <- fit |> reconcile(ols = min_trace(base, method = "ols"))
```


```{r}
fc |>
  filter(!is_aggregated(Hospital_ItemName_ENG), !is_aggregated(Grouped_Age)) |> 
  autoplot(data1_full) +
  facet_wrap(vars(Grouped_Organisation), scales = "free_y")+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```



```{r}
latest_date <- max(data1_full$YearMonth)  # Get the latest date
start_date <- latest_date - yearmonth(12)  # Subtract 12 months from the latest date

fc %>%
  filter(is_aggregated(Hospital_ItemName_ENG), is_aggregated(Grouped_Age)) %>%
  autoplot(data1_full) +
  facet_wrap(vars(Organisation), scales = "free_y") +
  coord_cartesian(xlim = c(start_date, latest_date))

```




```{r}
library(forecast)


# Convert data1_full to a time series object
ts_data <- ts(data1_full$Number, frequency = 12)  # Adjust frequency as needed

# Define forecast horizon
horizon <- 6 

# Fit an appropriate forecasting model
fit <- ets(ts_data)  

# Generate forecasts
forecast_result <- forecast(fit, h = horizon)

# Print the forecast results
print(forecast_result)

# Plot the forecast
plot(forecast_result)

```

```{r}

# Define forecast horizon
horizon <- 6  

# ---- Exponential Smoothing (ETS) Models ----

# Simple Exponential Smoothing (SES)
ses_fit <- ses(ts_data)
ses_forecast <- forecast(ses_fit, h = horizon)

# Holt's Linear Exponential Smoothing (Holt's Linear)
holt_fit <- holt(ts_data)
holt_forecast <- forecast(holt_fit, h = horizon)

# Holt-Winters Exponential Smoothing (Holt-Winters)
hw_fit <- hw(ts_data)
hw_forecast <- forecast(hw_fit, h = horizon)

# ---- ARIMA Models ----

# AutoRegressive Integrated Moving Average (ARIMA)
arima_fit <- auto.arima(ts_data)
arima_forecast <- forecast(arima_fit, h = horizon)

# Seasonal ARIMA (SARIMA)
sarima_fit <- auto.arima(ts_data, seasonal = TRUE)
sarima_forecast <- forecast(sarima_fit, h = horizon)



# ---- Plot Forecasts ----

# Plot ETS forecasts
plot(ses_forecast, main = "SES Forecast")
plot(holt_forecast, main = "Holt's Linear Forecast")
plot(hw_forecast, main = "Holt-Winters Forecast")

# Plot ARIMA forecasts
plot(arima_forecast, main = "ARIMA Forecast")
plot(sarima_forecast, main = "SARIMA Forecast")

```
# I am a bit confused on the forecast result. What are x-axis and y axis supposed to be?


-------------------------

```{r}
library(tidyverse)
library(hts)
library(readr)
library(furrr)
library(fabletools)
library(tscount)
```


```{r}
# Test sets of size 100,
origins <- 50 * seq(10) + 50
for (i in seq(origins)) {
  # Set up training set
  train <- data1
  train$bts <- subset(train$bts, end = nrow(data1$bts) - origins[i])
  # Create reconciled sample paths for different models
  reconcile_sample_paths(train, model_function = "ets")
  reconcile_sample_paths(train, model_function = "tscount")
  reconcile_sample_paths(train, model_function = "iglm")
  reconcile_sample_paths(train, model_function = "naiveecdf")
}

create_ensembles() # Include naive
create_ensembles(models_to_use = c("ets","iglm","tscount")) #Don't include naive
create_qcomb() # Include naive
create_qcomb(models_to_use = c("ets","iglm","tscount")) #Don't include naive

```

