---
title: "Hierarchical_2"
author: "Janice Hsu"
date: "2023-09-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(fpp3)
library(hts)
library(dplyr)
library(tidyr)
```


```{r}
data <- read.csv("HLTH0037_ts_cleaned.csv")
```


```{r}
data1 <- data %>%
  mutate(YearMonth = yearmonth(YearMonth)) %>%
  as_tsibble(index = YearMonth, key = c(Age_Code, Sex_ItemName_ENG, Hospital_Code, Hospital_ItemName_ENG)) 

data1 <- data1 %>%
  mutate(Number = 1)
```

#Number of patients entering ED under different hospital hierarchy
```{r}

data1_hts <- data1 %>%
  aggregate_key(Organisation/Hospital_ItemName_ENG, Number = sum(Number))
  
data1_hts |>
  filter(is_aggregated(Hospital_ItemName_ENG)) |>
  autoplot(Number) +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

•	A couple of Local Health Boards (LHBs) were redefined from the 1st of April 2019 onwards: Cwm Taf (27)--> Cwm Taf Morgannwg (30)// Abertawe Bro Morgannwg (26) --> Swansea Bay (31). Therefore, if you decide to forecast at LHB resolution, you might want to consider these 4 as a unique one. 
•	A the Princess of Wales Hospital changed its Local Health Boards
•	So we analyse these 4 as one organisation

# Group the changed Local Health Board together

```{r}

data1_grouped <- data1 %>%
  mutate(Grouped_Organisation = case_when(
    Organisation %in% c("Cwm Taf", "Cwm Taf Morgannwg", "Abertawe Bro Morgannwg", "Swansea Bay") ~ "Grouped_organisation",
    TRUE ~ Organisation
  ))

```

# There are 6 Local Health Boards
```{r}
unique(data1_grouped$Grouped_Organisation)
```

```{r}
data2_hts <- data1_grouped %>%
  group_by(Grouped_Organisation) %>%
  summarise(Number = sum(Number))
```

# Number of patients who enter ED under 6 different local health boards
```{r}
data2_hts |>
  ggplot(aes(x = YearMonth, y = Number)) +
  geom_line(stat = "identity") +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Grouped_Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


# Change the Age_Code structure into different groups

```{r}
unique(data1_grouped$Age_Code)
```


# Age group: "0-4", "5-17", "18-69", "70^"
```{r}
data1_grouped_age <- data1_grouped %>%
  filter(Age_Code != "Unknown") %>%
  mutate(Grouped_Age = case_when(
    Age_Code == "0 to 4" ~ "0-4",
    Age_Code == "5 to 17" ~ "5-17",
    Age_Code %in% c("18 to 24", "25 to 29", "30 to 34", "35 to 39", 
                    "40 to 44", "45 to 49", "50 to 54", "55 to 59", 
                    "60 to 64", "65 to 69") ~ "18-69",
    Age_Code %in% c("70 to 74", "75 to 79", "80 to 84", "85") ~ "70 and over",
    TRUE ~ "Other"
  ))

```


# Plot Number of Patients in different age groups
```{r}
data1_gts <- data1_grouped_age %>%
  filter(Sex_ItemName_ENG != "Not Specified or invalid") %>%
  group_by(Grouped_Age, Sex_ItemName_ENG) %>%
  summarize(Number = sum(Number, na.rm = TRUE))

ggplot(data1_gts, aes(x = Grouped_Age, y = Number)) +
  geom_bar(stat = "identity") +
  labs(y = "Number of patients") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Sex_ItemName_ENG, scales = "free")


```


#Change the data into wide format

```{r}

data2_wide <- data1_grouped %>%
  group_by(Grouped_Organisation) %>%
  index_by(YearMonth) %>%
  summarise(Number = sum(Number)) %>%
  pivot_wider(names_from = Grouped_Organisation, values_from = Number)

```


```{r}

```

-------------
# Forecast under Total (hierarchy 1) and Local Health Board (Hierarchy 2)



```{r}
# Create the hierarchical time series object
hts_data <- hts(data2_wide[,-1])  # exclude the YearMonth column

# Forecast using the hierarchical model
forecasts_hts <- forecast(hts_data)

# Plot the forecasts
plot(forecasts_hts)

# If you want to inspect the forecasts for the aggregated level:
aggregated_forecast <- forecasts_hts$allfcasts$Total

# If you want to inspect the forecasts for an individual health board, say "Aneurin Bevan":
aneurin_bevan_forecast <- forecasts_hts$allfcasts$`Aneurin Bevan`
```



```{r}
str(forecasts_hts)

```
* bts: This contains the forecasts for the different series.It's a time-series matrix.
* histy: This represents the historical data for the different series.
* labels: This list contains labels for the different levels in the hierarchy.
* method, fmethod: These are metadata about the forecasting process, indicating the method used for forecasting and combination.
* nodes: Information about the nodes at different levels in the hierarchy.

```{r}
# Required Libraries
library(zoo)
library(ggplot2)
library(lubridate)
library(tsibble)

# Dates for Historical Data
start_date <- yearmonth("2012 Apr")
end_date <- yearmonth("2023 Jun")
hist_dates <- seq(as.Date(start_date), as.Date(end_date), by = "1 month") %>% yearmonth()

# Extract historical time series data from the hts object
historical_data <- zoo(forecasts_hts$histy, order.by = hist_dates)

# Adjusting the forecast start date
forecast_start_date <- end_date + 1

# The rest of the code remains the same
forecast_dates <- seq(as.Date(forecast_start_date), by = "1 month", length.out = 10) %>% yearmonth()

# Extract forecast data from the hts object
forecast_data <- zoo(forecasts_hts$bts, order.by = forecast_dates)

# Combine historical and forecast data for plotting
combined_data <- rbind(historical_data, forecast_data)

# Convert the combined data to a long data frame for ggplot
df <- as.data.frame(fortify(combined_data, melt = TRUE))
names(df) <- c("Date", "Health_Board", "Number")

# Plot
ggplot(df, aes(x = Date, y = Number, color = Health_Board)) +
  geom_line() +
  labs(title = "Hierarchical Time Series Forecast",
       x = "YearMonth",
       y = "Number",
       color = "Health Board") +
  theme_minimal()

```


```{r}
print(combined_data)
```







