---
title: "Hierarchical_2"
author: "Janice Hsu"
date: "2023-09-15"
output: 
  bookdown::pdf_document2: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.pos = "H", out.extra = "")
```

\newpage

# Preliminary Analysis

```{r}
# Required Libraries
library(zoo)
library(ggplot2)
library(lubridate)
library(tsibble)
library(tidyverse)
library(fpp3)
library(hts)
library(dplyr)
library(tidyr)
library(forecast)
library(Metrics)
library(purrr)
library(tidyr)
library(stats)

```


```{r}
data <- read.csv("HLTH0037_ts_cleaned.csv")
```


## Data Introduction

The dataset contains 10 variables related to the hospitals and information of patients in Wales, UK. Here are the brief summary of the dataset:

- Data: This column represents the number of attendance in each emergency department.

- YearMonth: This column represents dates in the year-month format. Additionally, this dataset contains data from 2012 April to 2023 May.

- Age_Code: This column provides the age group that the patient is in. There are 17 different age groups. They are "0 to 4","5 to 17","18 to 24","25 to 29","30 to 34","35 to 39","40 to 44","45 to 49","50 to 54","55 to 59","60 to 64","65 to 69","70 to 74","75 to 79","80 to 84","85" and "Unknown".

- Sex_ItemName_ENG: This column provides the information of patient's gender. 

- Hospital_Code: This column represents 42 different hospitals in Wales.

- Hospital_ItemName_ENG: This columns refers to the name of the 42 different hospitals in Wales.

- Hospital_Hierarchy: This column represents the code for the health board that the hospital belongs to.

- Hospital_AltCode1: This column provides an alternate code for the hospital.

- Organisation: This column represents the health board.

- Organisation_Code: A code for the organisation as well as the health board.

- There are three hierarchies in this dataset. On the top level, there is all the hospitals in Wales, while on the second hierarchy, there are 6 different health boards which also shown as the organisations. At the bottom level, there are 42 hospitals in total.



\newpage
# Exploratory Data Analysis

```{r}
data <- data %>%
  mutate(YearMonth = yearmonth(YearMonth)) %>%
  as_tsibble(index = YearMonth, key = c(Age_Code, Sex_ItemName_ENG, Hospital_Code, Hospital_ItemName_ENG)) 

```

## Number of patients entering ED under different hospital hierarchy
```{r}

data_hts <- data %>%
  aggregate_key(Organisation/Hospital_ItemName_ENG, Number = sum(Data))
  
data_hts |>
  filter(is_aggregated(Hospital_ItemName_ENG)) |>
  autoplot(Number) +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1)) 
```

-	A couple of Local Health Boards (LHBs) were redefined from the 1st of April 2019 onwards: Cwm Taf (27)--> Cwm Taf Morgannwg (30)// Abertawe Bro Morgannwg (26) --> Swansea Bay (31). Therefore, if you decide to forecast at LHB resolution, you might want to consider these 4 as a unique one. 

-	A the Princess of Wales Hospital changed its Local Health Boards

-	So we analyse these 4 as one organisation

\newpage

## Group the changed Local Health Board together

```{r}

data_grouped <- data %>%
  mutate(Aggregated_Organisation = case_when(
    Organisation %in% c("Cwm Taf", "Cwm Taf Morgannwg", "Abertawe Bro Morgannwg", "Swansea Bay") ~ "Grouped_4_organisation",
    TRUE ~ Organisation
  ))

```

### There are 6 Local Health Boards
```{r}
unique(data_grouped$Aggregated_Organisation)
```

```{r}
data2_hts <- data_grouped %>%
  group_by(Aggregated_Organisation) %>%
  summarise(Number = sum(Data))
```

## Number of patients who enter ED under 6 different local health boards
```{r}
data2_hts |>
  ggplot(aes(x = YearMonth, y = Number)) +
  geom_line(stat = "identity") +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Aggregated_Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


Findings: 

- There is a big decline during the Covid-19 period, and after the Covid-19, except the Powys Teaching, other local health boards have increased the number of attendance back to its previous years.

- There seems to be seasonality in the data for each health board, and I would like to investigate more on top of this.

\newpage

## Seasonality of number of attendances

To investigate deeper into the potential seasonality in the data, I would like to decompose the time series according the each health board. It allows us to discover the trend, seasonality and the residual components.



### Decompose Time Series

```{r}

# Decompose time series for each health board using STL
stl_decompositions <- data2_hts %>%
  split(.$Aggregated_Organisation) %>%
  purrr::map(function(data){
    ts_data <- ts(data$Number, frequency = 12)
    stl(ts_data, s.window = "periodic")
  })

# Convert the decompositions to a tidy data frame
stl_df <- purrr::map2_dfr(stl_decompositions, names(stl_decompositions), 
                          ~{
                            time_series <- as.data.frame(.x$time.series)
                            time_series$Aggregated_Organisation <- .y
                            return(time_series)
                          })

```


```{r}
num_months <- length(unique(data2_hts$YearMonth))
num_orgs <- length(unique(data2_hts$Aggregated_Organisation))

# Correct assignment for YearMonth column
stl_df$YearMonth <- rep(unique(data2_hts$YearMonth), times = num_orgs)

```

### Plotting

```{r}
stl_df_tidy <- stl_df %>%
  tidyr::pivot_longer(cols = c(trend, seasonal, remainder), 
                      names_to = "component", 
                      values_to = "value")

```


```{r}
stl_df_tidy %>%
  ggplot(aes(x = YearMonth, y = value, color = component)) +
  geom_line() +
  facet_wrap(vars(Aggregated_Organisation), scales = "free_y", ncol = 3) +
  labs(title = "STL Decomposition of ED Attendance",
       y = "Number of Patients",
       color = "Component") +
  scale_color_manual(values = c("blue", "green", "red"),
                     breaks = c("trend", "seasonal", "remainder"),
                     labels = c("Trend", "Seasonal", "Residual")) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))

```
Based on the STL decomposition, it is evident that each health board follows a similar trend. Moreover, it is obvious that there is a significant downturn during the Covid-19 era. As for seasonality, there is a pronounced surge in the number of patient's attendance in the middle of the year (approximately in June or July). This seasoanl pattern underlines the recurrent nature of patient admissions.


\newpage

## Change the Age_Code structure into different groups

```{r}
unique(data_grouped$Age_Code)
```


### Age group: "0-4", "5-17", "18-69", "70^"
```{r}
data_grouped_age <- data_grouped %>%
  filter(Age_Code != "Unknown") %>%
  mutate(Grouped_Age = case_when(
    Age_Code == "0 to 4" ~ "0-4",
    Age_Code == "5 to 17" ~ "5-17",
    Age_Code %in% c("18 to 24", "25 to 29", "30 to 34", "35 to 39", 
                    "40 to 44", "45 to 49", "50 to 54", "55 to 59", 
                    "60 to 64", "65 to 69") ~ "18-69",
    Age_Code %in% c("70 to 74", "75 to 79", "80 to 84", "85") ~ "70 and over",
    TRUE ~ "Other"
  ))

```


## Plot Number of Patients in different age groups
```{r}
data_gts <- data_grouped_age %>%
  filter(Sex_ItemName_ENG != "Not Specified or invalid") %>%
  group_by(Grouped_Age, Sex_ItemName_ENG) %>%
  summarize(Number = sum(Data, na.rm = TRUE))

ggplot(data_gts, aes(x = Grouped_Age, y = Number)) +
  geom_bar(stat = "identity") +
  labs(y = "Number of patients") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~ Sex_ItemName_ENG, scales = "free")


```


Findings:

- The observation that the age group 18-69 has the most amount of patient attendance is expected, as it is the biggest group among all. However, it is noteworthy that the second biggest group are from the oldest age bracket, aligning with the general understanding of the health care need for the elders.


```{r}
#Change the data into wide format
data2_wide <- data_grouped %>%
  group_by(Aggregated_Organisation) %>%
  index_by(YearMonth) %>%
  summarise(Number = sum(Data)) %>%
  pivot_wider(names_from = Aggregated_Organisation, values_from = Number)


```

```{r}
data2_wide <- as_tibble(data2_wide)
data2_wide <- data2_wide %>%
   mutate(Total = rowSums(select(., c("Aneurin Bevan", "Betsi Cadwaladr", "Cardiff & Vale", "Grouped_4_organisation", "Hywel Dda", "Powys Teaching")), na.rm = TRUE))

```

# Forecast


```{r}
library(forecast)

# Convert data into time series
start_year <- year(min(data2_wide$YearMonth))
start_month <- month(min(data2_wide$YearMonth))

# Convert each column to a ts object
cols_to_forecast <- c("Aneurin Bevan", "Betsi Cadwaladr", "Cardiff & Vale", "Grouped_4_organisation", "Hywel Dda", "Powys Teaching", "Total")
forecast_list <- list()

for (col in cols_to_forecast) {
  ts_data <- ts(data2_wide[[col]], start = c(start_year, start_month), frequency = 12)
  
  # Forecast using auto.arima
  forecast_model <- auto.arima(ts_data)
  forecast_list[[col]] <- forecast(forecast_model, h = 6)
}

```

```{r}
print(forecast_list)

```



```{r}

h_forecast <- 12 
h_validate <- 6

# Splitting data
training_data <- head(data2_wide, nrow(data2_wide) - h_validate)
test_data <- tail(data2_wide, h_validate)


```



# Forecast with ARIMA

### Define the forecast horizon and validation period

```{r}
h <- 12
validation_period <- 6

```


### Splitting the data into training and testing sets
```{r}

training_data <- head(data2_wide, nrow(data2_wide) - validation_period)
test_data <- tail(data2_wide, validation_period)

```


```{r}
arima_forecast_list <- list()
start_year <- year(min(training_data$YearMonth))
start_month <- month(min(training_data$YearMonth))

for (col in cols_to_forecast) {
  ts_data <- ts(training_data[[col]], start = c(start_year, start_month), frequency = 12)
  arima_model <- auto.arima(ts_data)
  forecast_list[[col]] <- forecast(arima_model, h = h)
}
```

## Plotting using ARIMA

```{r}

for (col in cols_to_forecast) {
  end_year_train <- year(max(training_data$YearMonth))
  end_month_train <- month(max(training_data$YearMonth))
  
  start_year_test <- ifelse(end_month_train == 12, end_year_train + 1, end_year_train)
  start_month_test <- ifelse(end_month_train == 12, 1, end_month_train + 1)
  
  actual_ts <- ts(test_data[[col]], start = c(start_year_test, start_month_test), frequency = 12)
  
  plot_forecast <- autoplot(forecast_list[[col]]) +
    autolayer(actual_ts, series="Actual", PI=FALSE) +
    labs(title = paste("Forecast vs Actual for", col)) +
    theme(legend.position = "bottom")
  
  print(plot_forecast)
}
```




# Forecasting with ETS

```{r}

ets_forecast_list <- list()
start_year <- year(min(training_data$YearMonth))
start_month <- month(min(training_data$YearMonth))

for (col in cols_to_forecast) {
  ts_data <- ts(training_data[[col]], start = c(start_year, start_month), frequency = 12)
  ets_model <- ets(ts_data)  # Use the ets function
  forecast_list[[col]] <- forecast(ets_model, h = h)
}


```

## Plotting with ets
```{r}
for (col in cols_to_forecast) {
  end_year_train <- year(max(training_data$YearMonth))
  end_month_train <- month(max(training_data$YearMonth))
  
  start_year_test <- ifelse(end_month_train == 12, end_year_train + 1, end_year_train)
  start_month_test <- ifelse(end_month_train == 12, 1, end_month_train + 1)
  
  actual_ts <- ts(test_data[[col]], start = c(start_year_test, start_month_test), frequency = 12)
  
  plot_forecast <- autoplot(forecast_list[[col]]) +
    autolayer(actual_ts, series="Actual", PI=FALSE) +
    labs(title = paste("ETS Forecast vs Actual for", col)) +
    theme(legend.position = "bottom")
  
  print(plot_forecast)
}

```

# Accuracy assessment for ARIMA and ETS

```{r}
arima_error_metrics <- list()
ets_error_metrics <- list()

for (col in cols_to_forecast) {
  actual <- test_data[[col]]
  arima_forecasted <- head(forecast_list[[col]]$mean, validation_period)
  ets_forecasted <- head(ets_forecast_list[[col]]$mean, validation_period)
  
  # ARIMA metrics
  arima_MAE <- mae(actual, arima_forecasted)
  arima_RMSE <- rmse(actual, arima_forecasted)
  arima_MAPE <- tryCatch(mape(actual, arima_forecasted), error = function(e) NA)
  arima_error_metrics[[col]] <- list(MAE = arima_MAE, RMSE = arima_RMSE, MAPE = arima_MAPE)
  
  # ETS metrics
  ets_MAE <- mae(actual, ets_forecasted)
  ets_RMSE <- rmse(actual, ets_forecasted)
  ets_MAPE <- tryCatch(mape(actual, ets_forecasted), error = function(e) NA)
  ets_error_metrics[[col]] <- list(MAE = ets_MAE, RMSE = ets_RMSE, MAPE = ets_MAPE)
}

```


```{r}
for (col in cols_to_forecast) {
  cat("\nAccuracy metrics for:", col, "\n")
  cat("--------------------------\n")
  cat("ARIMA:\n")
  print(arima_error_metrics[[col]])
  cat("ETS:\n")
  print(ets_error_metrics[[col]])
  cat("\n")
}

```


