---
title: "Hierarchical Forecasting of Emergency Department Attendance"
subtitle: "<br/> with NHS"
author: "Janice Hsu"
institute: "Monash University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false



---
class: inverse, center, middle

# Introduction:

- Collaborate with NHS in the UK
- Forecast



---
class: inverse, center, middle

# Base Forecasting vs. Hierarchical Forecasting

### Different types of time series data
---
# Hierchical Structure

```{r setup, include=FALSE}
# Required Libraries
library(zoo)
library(ggplot2)
library(lubridate)
library(tsibble)
library(tidyverse)
library(fpp3)
library(hts)
library(dplyr)
library(tidyr)
library(forecast)
library(Metrics)
library(purrr)
library(tidyr)
library(stats)
library(fable)
library(dplyr)
library(igraph)
library(ggraph)
library(patchwork)
library(gridExtra)

```

```{r data, include=FALSE}
data <- read.csv("HLTH0037_ts_cleaned.csv")
```


```{r include=FALSE}
# change data structure
data <- data %>%
  mutate(YearMonth = yearmonth(YearMonth)) %>%
  as_tsibble(index = YearMonth, key = c(Age_Code, Sex_ItemName_ENG, Hospital_Code, Hospital_ItemName_ENG)) 

```


```{r small-plot, echo=FALSE, message=FALSE, warning=FALSE, out.width='80%',out.height='80%', fig.align='center', class.output='center-plot'}
library(igraph)

# Nodes data
nodes_df <- data.frame(
  name = c("All Wales", 
           "Betsi Cadwaladr", "Hywel Dda", "Grouped_4_organisation", 
           "Cardiff & Vale", "Aneurin Bevan", "Powys Teaching", 
           "Ysbyty Glan Clwyd", "Wrexham Maelor Hospital", "Colwyn Bay Community Hospital",
           "Holywell Community Hospital", "Mold Community Hospital"),
  stringsAsFactors = FALSE
)

# Edges data
edges_df <- data.frame(
  from = c(rep("All Wales", 6), 
           rep("Betsi Cadwaladr", 5)),
  to = c("Betsi Cadwaladr", "Hywel Dda", "Grouped_4_organisation", 
         "Cardiff & Vale", "Aneurin Bevan", "Powys Teaching",
         "Ysbyty Glan Clwyd", "Wrexham Maelor Hospital", "Colwyn Bay Community Hospital", 
         "Holywell Community Hospital", "Mold Community Hospital"),
  stringsAsFactors = FALSE
)


graph <- graph_from_data_frame(d = edges_df, vertices = nodes_df, directed = TRUE)


vertex_colors <- c("skyblue", rep("lightgreen", 6), rep("lightpink", 5))


# Plot the graph
plot(graph,
     vertex.size = 15,
     vertex.label = NA,  # Do not add labels
     vertex.color = vertex_colors,
     edge.arrow.size = 0.5,
     layout = layout.reingold.tilford(graph, root = 1))
legend("bottomright",
       legend = V(graph)$name,
       fill = vertex_colors,
       cex = 0.5,
       title = "Nodes",
       box.lty = 0)


```

---


# Number of patients entering ED under different hospital hierarchy
```{r attendance1, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%',out.height='60%', fig.align='center', class.output='center-plot'}
# Aggregate the data
data_hts <- data %>%
  aggregate_key(Organisation/Hospital_ItemName_ENG, attendance = sum(Data))

# Plot the aggregated data

data_hts |>
  filter(is_aggregated(Hospital_ItemName_ENG)) |>
  autoplot(attendance) +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1)) 
```
---
```{r data_grouped, echo=FALSE, warning=FALSE}
# mutate Aggregated_Organisation due to the change of the health boards

data_grouped <- data %>%
  mutate(Aggregated_Organisation = case_when(
    Organisation %in% c("Cwm Taf", "Cwm Taf Morgannwg", "Abertawe Bro Morgannwg", "Swansea Bay") ~ "Grouped_4_organisation",
    TRUE ~ Organisation
  ))

```


## There are 6 Local Health Boards

- Aneurin Bevan
- Betsi Cadwaladr
- Hywel Dda
- Grouped_4_organisation
- Cardiff & Vale
- Powys Teaching
---

```{r hts, echo = FALSE, warning=FALSE}
data2_hts <- data_grouped %>%
  group_by(Aggregated_Organisation) %>%
  summarise(attendance = sum(Data)) 
```

# Number of patients who enter ED under 6 different local health boards
```{r attendance2, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='60%', fig.align='center', class.output='center-plot'}
data2_hts |>
  ggplot(aes(x = YearMonth, y = attendance)) +
  geom_line(stat = "identity") +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Aggregated_Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
---
# Findings:
- COVID-19 caused significant changes in patient attendance.

- Due to health regulations and concerns about virus infection, attendance at many local health boards decreased during the pandemic.

- The post-pandemic period showed a rise in attendance, with various boards showing distinct recovery path.

- Powys Teaching noticeably departed from the general upward recovery trend.

- Data show that attendance patterns appear to be seasonal across all health boards.

- Future efforts should concentrate on determining this seasonality's sources, effects, and incorporation into predictive forecasting models.

---

# Seasonality of number of attendances

```{r stl, echo=FALSE, warning=FALSE}

# Decompose time series for each health board using STL
stl_decompositions <- data2_hts %>%
  split(.$Aggregated_Organisation) %>%
  purrr::map(function(data){
    ts_data <- ts(data$attendance, frequency = 12)
    stl(ts_data, s.window = "periodic")
  })

# Convert the decompositions to a tidy data frame
stl_df <- purrr::map2_dfr(stl_decompositions, names(stl_decompositions), 
                          ~{
                            time_series <- as.data.frame(.x$time.series)
                            time_series$Aggregated_Organisation <- .y
                            return(time_series)
                          })

```


```{r length, echo=FALSE, warning=FALSE}
num_months <- length(unique(data2_hts$YearMonth))
num_orgs <- length(unique(data2_hts$Aggregated_Organisation))

# Correct assignment for YearMonth column
stl_df$YearMonth <- rep(unique(data2_hts$YearMonth), times = num_orgs)

```

### Plotting

```{r longer, echo=FALSE, warning=FALSE}
stl_df_tidy <- stl_df %>%
  tidyr::pivot_longer(cols = c(trend, seasonal, remainder), 
                      names_to = "component", 
                      values_to = "value")

```


```{r stl_decom, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='60%', fig.align='center', class.output='center-plot'}
stl_df_tidy %>%
  ggplot(aes(x = YearMonth, y = value, color = component)) +
  geom_line() +
  facet_wrap(vars(Aggregated_Organisation), scales = "free_y", ncol = 3) +
  labs(title = "STL Decomposition of ED Attendance",
       y = "Number of Patients",
       color = "Component") +
  scale_color_manual(values = c("blue", "green", "red"),
                     breaks = c("trend", "seasonal", "remainder"),
                     labels = c("Trend", "Seasonal", "Residual")) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))

```



---
class: center, middle

- **Similar trends:** Consistent patient attendance patterns are revealed by STL decomposition across all health boards.
  
- **Impact of COVID-19:** The Covid-19 era is marked by a noticeable drop in attendance, which implies to a significant impact on healthcare utilization or access.
  
- **An increase in mid-year attendance:** Every year, there is a noticeable rise in patient attendance around June or July, according to seasonal variations.
  
---

```{r  echo=FALSE, warning=FALSE}
#Change the data into wide format
data2_wide <- data_grouped %>%
  group_by(Aggregated_Organisation) %>%
  index_by(YearMonth) %>%
  summarise(attendance = sum(Data)) %>%
  pivot_wider(names_from = Aggregated_Organisation, values_from = attendance)


```

```{r  echo=FALSE, warning=FALSE}
data2_wide <- as_tibble(data2_wide)
data2_wide <- data2_wide %>%
   mutate(Total = rowSums(select(., c("Aneurin Bevan", "Betsi Cadwaladr", "Cardiff & Vale", "Grouped_4_organisation", "Hywel Dda", "Powys Teaching")), na.rm = TRUE))

```



# Forecast with ARIMA



```{r  echo=FALSE, warning=FALSE}

# Define the forecast horizon and validation period
h <- 12
validation_period <- 6

```



```{r  echo=FALSE, warning=FALSE}
# Splitting the data into training and testing sets
training_data <- head(data2_wide, nrow(data2_wide) - validation_period)
test_data <- tail(data2_wide, validation_period)

```


```{r  echo=FALSE, warning=FALSE}

cols_to_forecast <- c("Aneurin Bevan", "Betsi Cadwaladr", "Cardiff & Vale", "Grouped_4_organisation", "Hywel Dda", "Powys Teaching", "Total")
forecast_list <- list()

arima_forecast_list <- list()
start_year <- year(min(training_data$YearMonth))
start_month <- month(min(training_data$YearMonth))


for (col in cols_to_forecast) {
  ts_data <- ts(training_data[[col]], start = c(start_year, start_month), frequency = 12)
  arima_model <- auto.arima(ts_data)
  forecast_list[[col]] <- forecast(arima_model, h = h)
}
```



```{r  echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot', fig.show='asis'}

arima_plots <- list()
for (col in cols_to_forecast) {
  
  end_year_train <- year(max(training_data$YearMonth))
  end_month_train <- month(max(training_data$YearMonth))
  
  start_year_test <- ifelse(end_month_train == 12, end_year_train + 1, end_year_train)
  start_month_test <- ifelse(end_month_train == 12, 1, end_month_train + 1)
  
  actual_ts <- ts(test_data[[col]], start = c(start_year_test, start_month_test), frequency = 12)
  
   p <- autoplot(forecast_list[[col]]) + 
        autolayer(actual_ts, series='Actual', PI=FALSE) + 
        labs(title = paste('Forecast vs Actual for', col)) + 
        theme(legend.position = 'bottom')
    
    arima_plots[[col]] <- p
}



```







```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Total']])
```
---

# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Aneurin Bevan']])
```



---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Betsi Cadwaladr']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Cardiff & Vale']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Grouped_4_organisation']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Hywel Dda']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Hywel Dda']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Powys Teaching']])
```


---
# Forecasting with ETS

```{r echo = FALSE}

ets_forecast_list <- list()
start_year <- year(min(training_data$YearMonth))
start_month <- month(min(training_data$YearMonth))

for (Aggregated_Organisation in cols_to_forecast) {
  ts_data <- ts(training_data[[Aggregated_Organisation]], start = c(start_year, start_month), frequency = 12)
  ets_model <- ets(ts_data) 
  forecast_list[[Aggregated_Organisation]] <- forecast(ets_model, h = h)
}


```



```{r  echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot', fig.show='asis'}

ets_plots <- list()
for (col in cols_to_forecast) {
  
  end_year_train <- year(max(training_data$YearMonth))
  end_month_train <- month(max(training_data$YearMonth))
  
  start_year_test <- ifelse(end_month_train == 12, end_year_train + 1, end_year_train)
  start_month_test <- ifelse(end_month_train == 12, 1, end_month_train + 1)
  
  actual_ts <- ts(test_data[[col]], start = c(start_year_test, start_month_test), frequency = 12)
  
   p <- autoplot(forecast_list[[col]]) + 
        autolayer(actual_ts, series='Actual', PI=FALSE) + 
        labs(title = paste('Forecast vs Actual for', col)) + 
        theme(legend.position = 'bottom')
    
    ets_plots[[col]] <- p
}



```



```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Total']])
```
---

# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Aneurin Bevan']])
```



---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Betsi Cadwaladr']])
```

---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Cardiff & Vale']])
```

---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Grouped_4_organisation']])
```

---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Hywel Dda']])
```


---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Powys Teaching']])
```

---

class: center, middle
**Result:**

- With Cardiff & Vale being a significant exception, most places displayed reduced error metrics when modeled with ETS, signifying a higher predictive accuracy as compared to the ARIMA model.



---

# Reconciliation

## Step 1: Aggregate the data


```{r echo=FALSE, warning=FALSE, include = FALSE}
data_grouped_age <- data_grouped %>%
  filter(Age_Code != "Unknown") %>%
  mutate(Grouped_Age = case_when(
    Age_Code == "0 to 4" ~ "0-4",
    Age_Code == "5 to 17" ~ "5-17",
    Age_Code %in% c("18 to 24", "25 to 29", "30 to 34", "35 to 39", 
                    "40 to 44", "45 to 49", "50 to 54", "55 to 59", 
                    "60 to 64", "65 to 69") ~ "18-69",
    Age_Code %in% c("70 to 74", "75 to 79", "80 to 84", "85") ~ "70 and over",
    TRUE ~ "Other"
  ))

```


```{r}
# hierarchy with age group
data_hts <- data_grouped_age |>
  aggregate_key(Aggregated_Organisation / Hospital_ItemName_ENG * Grouped_Age, attendance = sum(Data))
```


```{r}
# hierarchy without age group, only aggregate with the organisation, wales as a big picture
data_wales <- data_grouped_age |>
  aggregate_key(Aggregated_Organisation, attendance = sum(Data))
```

---

# Step 2: Time series cross validation

```{r echo=FALSE, warning=FALSE, include = FALSE}
library(tscount)
# Split data
training_data <- data_wales %>% 
  filter(YearMonth <= yearmonth("2022 Jun"))

test_data <- data_wales %>% 
  filter(YearMonth > yearmonth("2022 Jun")) 

test_data_filtered <- test_data %>%
  filter(Aggregated_Organisation == "<aggregated>")

```


```{r}
# average the three models
library(fable.tscount)
fit_cv<- training_data %>%
  model(
    naive_model = NAIVE(attendance),
    ets_model = ETS(attendance),
    tscount = TSCOUNT(attendance ~ trend() , link = "log", model = list(past_obs = 1:3))
  )|> mutate (comb = (naive_model+ets_model+tscount)/3)
```




```{r echo=FALSE, warning=FALSE}
# Generate forecasts
fc_cv <- fit_cv %>%
  forecast(h = 12) %>%
  filter(.model == "comb") 

```


```{r warning=FALSE}
# Generate forecasts for aggregated
fc_cv_filtered <- fit_cv %>%
  forecast(h = 12) %>%
  filter(.model == "comb") %>%
  filter(Aggregated_Organisation == "<aggregated>")
```

---

# Check the accuracy

```{r}

forecasted_values <- fc_cv_filtered$.mean

accuracy_value <- accuracy(
  as.data.frame(forecasted_values),as.data.frame(test_data_filtered$attendance))

print(accuracy_value)


```

---

```{r echo=FALSE, warning=FALSE}

joined_data <- as_tibble(test_data) %>%
  left_join(as_tibble(fc_cv), by = c("YearMonth", "Aggregated_Organisation")) %>%
  select(-attendance.y) %>%
  rename(attendance = attendance.x) %>%
  rename(forecast = .mean)

```


```{r echo=FALSE, warning=FALSE}

comparison_data <- data.frame(
  Month = yearmonth(test_data$YearMonth),
  Actual = test_data$attendance,
  Forecasted = fc_cv$.mean,
  Aggregated_Organisation = test_data$Aggregated_Organisation  
) %>%rowwise()

```


```{r echo=FALSE, message=FALSE, warning=FALSE}

joined_data_long<- joined_data %>%
  pivot_longer(
    cols = c(attendance, forecast),
    names_to = "Type",
    values_to = "Value"
  )



unique_organisations <- unique(joined_data_long$Aggregated_Organisation)

plot_list <- map(unique_organisations, ~{
  ggplot(joined_data_long[joined_data_long$Aggregated_Organisation == .x,], 
         aes(x = YearMonth, y = Value, color = Type)) +
    geom_line() +
    labs(title = paste("Forecasted vs Actual Values for", .x),
         y = "Values") +
    theme_minimal()
})



```



# Step 3: Forecast

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
plot_list[[1]]
```
---

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
plot_list[[2]]
```

---

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
plot_list[[3]]
```

---

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
plot_list[[4]]
```

---

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
plot_list[[5]]
```

---

```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
plot_list[[6]]
```

---
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
plot_list[[7]]
```

---
# Running t-tests and extracting p-value and t-value
```{r}

t_test_summary <- joined_data %>%
  summarise(
    t_value = list(t.test(attendance, forecast)$statistic),
    p_value = list(t.test(attendance, forecast)$p.value)
  ) %>%
  unnest(cols = c(t_value, p_value))

print(t_test_summary)

```

The p value is relatively high. It suggests that we fail to reject the null hypothesis, meaning there is no enough evidence to conclude that forecast and actual values are significant different. In other words, the t-value of 0.313 is not large enough to be considered statistically significant.

---

# Reconcile


```{r echo=FALSE, warning=FALSE}
# 1. Fitting models and Forecasting
fit_wales <- data_wales |>
  model(
    naive_model = NAIVE(attendance),
    ets_model = ETS(attendance),
    tscount = TSCOUNT(attendance ~ trend() , link = "log", model = list(past_obs = 1:3))
  )|> mutate (comb = (naive_model+ets_model+tscount)/3)
```
  
```{r echo=FALSE, warning=FALSE}
wales_rc <- fit_wales %>%
   reconcile(bu = bottom_up(comb),
            wls_ETS1 = min_trace(comb, method = "wls_struct"),
            wls_ETS2 = min_trace(comb, method = "wls_var"),
            mint_ETS = min_trace(comb, method = "mint_shrink")) %>%
  forecast(h = 6)
```



```{r echo=FALSE, warning=FALSE}
forecast_wales_avg <- wales_rc %>%
  as_tibble() %>%
  group_by(Aggregated_Organisation, YearMonth) %>%
  summarise(mean_attendance = mean(.mean, na.rm = TRUE), .groups = "drop")

```

```{r echo=FALSE, warning=FALSE}
forecast_wales_avg$Aggregated_Organisation <- factor(forecast_wales_avg$Aggregated_Organisation)

forecast_wales_avg <- forecast_wales_avg %>%
  rename(attendance = mean_attendance)

```


```{r echo=FALSE, warning=FALSE}

combined_data_rc <- bind_rows(data2_hts, forecast_wales_avg)

```


```{r echo=FALSE, warning=FALSE}

combined_data_rc <- combined_data_rc %>%
  mutate(data_type = if_else(YearMonth < ymd("2023-06-01"), "Historical", "Forecast"))


# Plotting
ggplot(combined_data_rc, aes(x = YearMonth, y = attendance, color = data_type)) +
  geom_line() +
  facet_wrap(~Aggregated_Organisation, scales = "free_y") +
  theme_minimal() +
  labs(title = "Attendance Over Time",
       subtitle = "Distinguishing between historical and forecasted data",
       x = "Year and Month",
       y = "Attendance",
       color = "Data Type") +
  scale_color_manual(values = c("Historical" = "blue", "Forecast" = "red"))

```




---
class: center, middle
# Conclusion

Why the Trend is Declining:

- External Factors: Societal changes, economic downturns, and pandemics can all have a big impact. For instance, the COVID-19 pandemic reduced hospital visits as a result of patients' concern over spreading the illness.

- Changes in healthcare legislation, policies, or insurance provisions may deter people from obtaining medical care unless it is absolutely essential.

- Seasonal Effects: Some months or seasons may naturally draw lesser crowds, maybe as a result of a lack of prevalent diseases or ailments that peak in other months.

