---
title: "Hierarchical Forecasting of Emergency Department Attendance"
subtitle: "⚔<br/> with NHS"
author: "Janice Hsu"
institute: "Monash University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false



---
class: inverse, center, middle

# Base Forecasting vs. Hierarchical Forecasting

### Different types of time series data

---

# Hierchical Structure

```{r setup, include=FALSE}
# Required Libraries
library(zoo)
library(ggplot2)
library(lubridate)
library(tsibble)
library(tidyverse)
library(fpp3)
library(hts)
library(dplyr)
library(tidyr)
library(forecast)
library(Metrics)
library(purrr)
library(tidyr)
library(stats)
library(fable)
library(dplyr)
library(igraph)
library(ggraph)
library(patchwork)
library(gridExtra)

```

```{r data, include=FALSE}
data <- read.csv("HLTH0037_ts_cleaned.csv")
```


```{r include=FALSE}
# change data structure
data <- data %>%
  mutate(YearMonth = yearmonth(YearMonth)) %>%
  as_tsibble(index = YearMonth, key = c(Age_Code, Sex_ItemName_ENG, Hospital_Code, Hospital_ItemName_ENG)) 

```


```{r small-plot, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%',out.height='60%', fig.align='center', class.output='center-plot'}
library(igraph)

# Nodes data
nodes_df <- data.frame(
  name = c("All Wales", 
           "Betsi Cadwaladr", "Hywel Dda", "Grouped_4_organisation", 
           "Cardiff & Vale", "Aneurin Bevan", "Powys Teaching", 
           "Ysbyty Glan Clwyd", "Wrexham Maelor Hospital", "Colwyn Bay Community Hospital", 
           "Holywell Community Hospital", "Mold Community Hospital"),
  stringsAsFactors = FALSE
)

# Edges data
edges_df <- data.frame(
  from = c(rep("All Wales", 6), 
           rep("Betsi Cadwaladr", 5)),
  to = c("Betsi Cadwaladr", "Hywel Dda", "Grouped_4_organisation", 
         "Cardiff & Vale", "Aneurin Bevan", "Powys Teaching",
         "Ysbyty Glan Clwyd", "Wrexham Maelor Hospital", "Colwyn Bay Community Hospital", 
         "Holywell Community Hospital", "Mold Community Hospital"),
  stringsAsFactors = FALSE
)

# Creating graph object
graph <- graph_from_data_frame(d = edges_df, vertices = nodes_df, directed = TRUE)

# Color vector: 1 color for "All Wales", 6 colors for organisations, and 5 colors for hospitals
vertex_colors <- c("skyblue", rep("lightgreen", 6), rep("lightpink", 5))

# Plot the graph
plot(graph, 
     vertex.size = 15, 
     vertex.label.cex = 0.6,  
     vertex.color = vertex_colors,  
     edge.arrow.size = 0.5, 
     vertex.label.degree = pi/4, # 45-degree label rotation
     layout = layout.reingold.tilford(graph, root = 1))

```

---

# Number of patients entering ED under different hospital hierarchy
```{r attendance1, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%',out.height='60%', fig.align='center', class.output='center-plot'}
# Aggregate the data
data_hts <- data %>%
  aggregate_key(Organisation/Hospital_ItemName_ENG, attendance = sum(Data))

# Plot the aggregated data

data_hts |>
  filter(is_aggregated(Hospital_ItemName_ENG)) |>
  autoplot(attendance) +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(angle = 0, hjust = 1)) 
```
---
```{r data_grouped, echo=FALSE, warning=FALSE}
# mutate Aggregated_Organisation due to the change of the health boards

data_grouped <- data %>%
  mutate(Aggregated_Organisation = case_when(
    Organisation %in% c("Cwm Taf", "Cwm Taf Morgannwg", "Abertawe Bro Morgannwg", "Swansea Bay") ~ "Grouped_4_organisation",
    TRUE ~ Organisation
  ))

```

class: center
## There are 6 Local Health Boards

- Aneurin Bevan
- Betsi Cadwaladr
- Hywel Dda
- Grouped_4_organisation
- Cardiff & Vale
- Powys Teaching
---

```{r hts, echo = FALSE, warning=FALSE}
data2_hts <- data_grouped %>%
  group_by(Aggregated_Organisation) %>%
  summarise(attendance = sum(Data)) 
```

# Number of patients who enter ED under 6 different local health boards
```{r attendance2, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='60%', fig.align='center', class.output='center-plot'}
data2_hts |>
  ggplot(aes(x = YearMonth, y = attendance)) +
  geom_line(stat = "identity") +
  labs(y = "Number of patients",
       title = "Number of patients who enter ED") +
  facet_wrap(vars(Aggregated_Organisation), scales = "free_y", ncol = 3) +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
---
# Findings:
- COVID-19 caused significant changes in patient attendance.

- Due to health regulations and concerns about virus infection, attendance at many local health boards decreased during the pandemic.

- The post-pandemic period showed a rise in attendance, with various boards showing distinct recovery path.

- Powys Teaching noticeably departed from the general upward recovery trend.

- Data show that attendance patterns appear to be seasonal across all health boards.

- Future efforts should concentrate on determining this seasonality's sources, effects, and incorporation into predictive forecasting models.

---

# Seasonality of number of attendances

```{r stl, echo=FALSE, warning=FALSE}

# Decompose time series for each health board using STL
stl_decompositions <- data2_hts %>%
  split(.$Aggregated_Organisation) %>%
  purrr::map(function(data){
    ts_data <- ts(data$attendance, frequency = 12)
    stl(ts_data, s.window = "periodic")
  })

# Convert the decompositions to a tidy data frame
stl_df <- purrr::map2_dfr(stl_decompositions, names(stl_decompositions), 
                          ~{
                            time_series <- as.data.frame(.x$time.series)
                            time_series$Aggregated_Organisation <- .y
                            return(time_series)
                          })

```


```{r length, echo=FALSE, warning=FALSE}
num_months <- length(unique(data2_hts$YearMonth))
num_orgs <- length(unique(data2_hts$Aggregated_Organisation))

# Correct assignment for YearMonth column
stl_df$YearMonth <- rep(unique(data2_hts$YearMonth), times = num_orgs)

```

### Plotting

```{r longer, echo=FALSE, warning=FALSE}
stl_df_tidy <- stl_df %>%
  tidyr::pivot_longer(cols = c(trend, seasonal, remainder), 
                      names_to = "component", 
                      values_to = "value")

```


```{r stl_decom, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='60%', fig.align='center', class.output='center-plot'}
stl_df_tidy %>%
  ggplot(aes(x = YearMonth, y = value, color = component)) +
  geom_line() +
  facet_wrap(vars(Aggregated_Organisation), scales = "free_y", ncol = 3) +
  labs(title = "STL Decomposition of ED Attendance",
       y = "Number of Patients",
       color = "Component") +
  scale_color_manual(values = c("blue", "green", "red"),
                     breaks = c("trend", "seasonal", "remainder"),
                     labels = c("Trend", "Seasonal", "Residual")) +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1))

```



---
class: center

- **Similar trends:** Consistent patient attendance patterns are revealed by STL decomposition across all health boards.
  
- **Impact of COVID-19:** The Covid-19 era is marked by a noticeable drop in attendance, which implies to a significant impact on healthcare utilization or access.
  
- **An increase in mid-year attendance:** Every year, there is a noticeable rise in patient attendance around June or July, according to seasonal variations.
  
---

```{r  echo=FALSE, warning=FALSE}
#Change the data into wide format
data2_wide <- data_grouped %>%
  group_by(Aggregated_Organisation) %>%
  index_by(YearMonth) %>%
  summarise(attendance = sum(Data)) %>%
  pivot_wider(names_from = Aggregated_Organisation, values_from = attendance)


```

```{r  echo=FALSE, warning=FALSE}
data2_wide <- as_tibble(data2_wide)
data2_wide <- data2_wide %>%
   mutate(Total = rowSums(select(., c("Aneurin Bevan", "Betsi Cadwaladr", "Cardiff & Vale", "Grouped_4_organisation", "Hywel Dda", "Powys Teaching")), na.rm = TRUE))

```



# Forecast with ARIMA



```{r  echo=FALSE, warning=FALSE}

# Define the forecast horizon and validation period
h <- 12
validation_period <- 6

```



```{r  echo=FALSE, warning=FALSE}
# Splitting the data into training and testing sets
training_data <- head(data2_wide, nrow(data2_wide) - validation_period)
test_data <- tail(data2_wide, validation_period)

```


```{r  echo=FALSE, warning=FALSE}

cols_to_forecast <- c("Aneurin Bevan", "Betsi Cadwaladr", "Cardiff & Vale", "Grouped_4_organisation", "Hywel Dda", "Powys Teaching", "Total")
forecast_list <- list()

arima_forecast_list <- list()
start_year <- year(min(training_data$YearMonth))
start_month <- month(min(training_data$YearMonth))


for (col in cols_to_forecast) {
  ts_data <- ts(training_data[[col]], start = c(start_year, start_month), frequency = 12)
  arima_model <- auto.arima(ts_data)
  forecast_list[[col]] <- forecast(arima_model, h = h)
}
```



```{r  echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot', fig.show='asis'}

arima_plots <- list()
for (col in cols_to_forecast) {
  
  end_year_train <- year(max(training_data$YearMonth))
  end_month_train <- month(max(training_data$YearMonth))
  
  start_year_test <- ifelse(end_month_train == 12, end_year_train + 1, end_year_train)
  start_month_test <- ifelse(end_month_train == 12, 1, end_month_train + 1)
  
  actual_ts <- ts(test_data[[col]], start = c(start_year_test, start_month_test), frequency = 12)
  
   p <- autoplot(forecast_list[[col]]) + 
        autolayer(actual_ts, series='Actual', PI=FALSE) + 
        labs(title = paste('Forecast vs Actual for', col)) + 
        theme(legend.position = 'bottom')
    
    arima_plots[[col]] <- p
}



```







```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Total']])
```
---

# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Aneurin Bevan']])
```



---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Betsi Cadwaladr']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Cardiff & Vale']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Grouped_4_organisation']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Hywel Dda']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Hywel Dda']])
```

---
# Arima model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(arima_plots[['Powys Teaching']])
```


---
# Forecasting with ETS

```{r echo = FALSE}

ets_forecast_list <- list()
start_year <- year(min(training_data$YearMonth))
start_month <- month(min(training_data$YearMonth))

for (Aggregated_Organisation in cols_to_forecast) {
  ts_data <- ts(training_data[[Aggregated_Organisation]], start = c(start_year, start_month), frequency = 12)
  ets_model <- ets(ts_data) 
  forecast_list[[Aggregated_Organisation]] <- forecast(ets_model, h = h)
}


```



```{r  echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot', fig.show='asis'}

ets_plots <- list()
for (col in cols_to_forecast) {
  
  end_year_train <- year(max(training_data$YearMonth))
  end_month_train <- month(max(training_data$YearMonth))
  
  start_year_test <- ifelse(end_month_train == 12, end_year_train + 1, end_year_train)
  start_month_test <- ifelse(end_month_train == 12, 1, end_month_train + 1)
  
  actual_ts <- ts(test_data[[col]], start = c(start_year_test, start_month_test), frequency = 12)
  
   p <- autoplot(forecast_list[[col]]) + 
        autolayer(actual_ts, series='Actual', PI=FALSE) + 
        labs(title = paste('Forecast vs Actual for', col)) + 
        theme(legend.position = 'bottom')
    
    ets_plots[[col]] <- p
}



```



# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Total']])
```
---

# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Aneurin Bevan']])
```



---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Betsi Cadwaladr']])
```

---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Cardiff & Vale']])
```

---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Grouped_4_organisation']])
```

---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Hywel Dda']])
```

---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Hywel Dda']])
```

---
# ETS model
```{r echo=FALSE, message=FALSE, warning=FALSE, out.width='70%', out.height='70%', fig.align='center', class.output='center-plot'}
print(ets_plots[['Powys Teaching']])
```

---

class: center
**Result:**

- With Cardiff & Vale being a significant exception, most places displayed reduced error metrics when modeled with ETS, signifying a higher predictive accuracy as compared to the ARIMA model.

---
class: center

- **Base forecasting:** applied when forecasting simple time series.

- **Reconciliation forecasting (Hierarchical forecasting):** it includes forecasting on different levels of aggregation, and it should be applied when dealing with hierarchical or grouped time series.

- **Notes:** RMSE and MAE should not be used for the hierarchical time series data (Problems with reconciliation, Metric incomparability)

---

# Reconciliation

## Step 1: Aggregate the data


```{r echo=FALSE, warning=FALSE}
data_grouped_age <- data_grouped %>%
  filter(Age_Code != "Unknown") %>%
  mutate(Grouped_Age = case_when(
    Age_Code == "0 to 4" ~ "0-4",
    Age_Code == "5 to 17" ~ "5-17",
    Age_Code %in% c("18 to 24", "25 to 29", "30 to 34", "35 to 39", 
                    "40 to 44", "45 to 49", "50 to 54", "55 to 59", 
                    "60 to 64", "65 to 69") ~ "18-69",
    Age_Code %in% c("70 to 74", "75 to 79", "80 to 84", "85") ~ "70 and over",
    TRUE ~ "Other"
  ))

```


```{r}
# hierarchy with age group
data_hts <- data_grouped_age |>
  aggregate_key(Aggregated_Organisation / Hospital_ItemName_ENG * Grouped_Age, attendance = sum(Data))
```


```{r}
# hierarchy without age group, only aggregate with the organisation, wales as a big picture
data_wales <- data_grouped_age |>
  aggregate_key(Aggregated_Organisation, attendance = sum(Data))
```



# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](https://yihui.org/knitr/), and [R Markdown](https://rmarkdown.rstudio.com).
